Bayesian methods for black hole merger parameter estimation with LISA.

### 2019-04-04 paper/paper_PE.tex

>>>>

For this exploratory work we take a ``naive'' approach to  estimating posterior-probability distributed parameter samples, not applying any specific knowledge or expectations about distributions which might plausibly lead to artifically robust-seeming but incomplete results.  We apply two independent approaches, an parallel tempering MCMC sampler with generically wtih amotivated proposal distribution, and nested sampling. \jgb{More to come}

-- Bayesian samplers: Multinest, PTMC

For comparison with our Bayesian inference results we also compute estimates of parameter estimation precision using the Fisher information matrix the common workhorse of LISA science studies despite its limitations.~\cite{Vallisneri08}  For measurements with additive Gaussian noise, as assumed here, the Fisher information matrix can be computed by
\be
F_{ij}=\left( \partial_{\theta_i} h| \partial_{\theta_i} h \right).
\ee
The inverse of this matrix estimates the covariance matrix describing parameter uncertainty. 
The derivatives are computed by second-order finite differences, with the signal inner-products expanded and computed before differencing. For these calculations, where numerical smoothness is more of a concern than speed we compute the inner products explicitly on a fine grid.

Generally, finite difference Fisher matrices can be problematic, with small numerical defects potentially having outsized impact, so the finite-difference step size $\epsilon_i$ used in each parameter derivative $\partial_{\theta_i}h$ must be chosen carefully. We target the step-size to be scaled off the diagonal Fisher matrix elements by ${\epsilon_i}\approx\delta/\sqrt{F_{ii}}$ using $\delta=0.001$ \jgb{This is \texttt{fisher\_err\_target}, was it different for some runs?}  for the results shown here.  To achieve this we set begin with an initial choice of the $\epsilon_i$ scaled of parameter values or prior widths, then we estimate the Fisher diagonals and iterate until convergence. In this process we also constrain that the step is never more than $10^{-10}$ times the initial scaling. \jgb{There are some more special-case details, but probably not worth elaborating.  We could also discuss how we do matrix inversion, but I don't think there is anything interesting there.}
